---
title: 'Reproducible Research: Creating a Table One'
author: "Mireya Vaca"
date: "January 16, 2015"
output: html_document
---


This file will go through the steps and the Rcode necessary for creating a Table One from data in the file "hw1data.csv" using an existing package.


First, read the data into R. 

```{r}
hw1data = read.csv("hw1data.csv")
```


Now, display the first few rows of the data to look at what we have. 

```{r}
head(hw1data)
```


You will need to load the package for creating a table into R.  The package can be downloaded from:  
http://cran.r-project.org/web/packages/tableone/index.html  

```{r}
library(tableone)
```


There are two categorical variables, sex and race. Categorical variables will need to be transformed into factors. 

```{r}
catvarstofactors <- c("sex", "race")
hw1data[catvarstofactors] <- lapply(hw1data[catvarstofactors], factor)
```


Obtain the names of the variables. 

```{r}
dput(names(hw1data))
```


Then, create a list of the variables that you want to summarize. This is a vector of variables.

```{r}
vars <-c ("age", "sex", "race", "weight", "height")
```


Now, create the table, stratifying by treatment group. 

```{r}
TableOne <- CreateTableOne(vars=vars, strata=("group"), data=hw1data)
```


When you print the table, you may want to show all levels included in the categorical variables.  

```{r}
print(TableOne, showAllLevels=TRUE)
```


Before reporting the results, you will want to run a summary of the data to check for normality of continuous variables, missing values, and frequency of data to make sure the right statistical tests were used. 

```{r}
summary(TableOne)
```


The summary statistics show that data for the continuous variables follow a normal distribution (based on skewness and mean/median data), there are no missing values, and frequencies in each category are large. So, we can use the default tests in the package. The tests used are chi-square tests with continuity correction for categorical variables and a two-sample t-test for continuous variables (under an equal variance assumption). 


This is the table we will use for Table One.

```{r}
print(TableOne, showAllLevels=TRUE)
```


Additional Resources on using the TableOne package:  
* http://www.inside-r.org/node/236481   
* http://rpubs.com/kaz_yos/tableone-vignette

