---
title: 'Reproducible Research: Creating a Table One'
author: "Mireya Vaca"
date: "January 16, 2015"
output: html_document
---


This file will go through the steps and the Rcode necessary for creating a Table One from data in the file "hw1data.csv" using an existing package.


First, read the data into R. 

```{r}
hw1data = read.csv("hw1data.csv")
```


Now, display the first few rows of the data to look at what we have. 

```{r}
head(hw1data)
```


You will need to load the package for creating a table into R.  The package can be downloaded from: http://cran.r-project.org/web/packages/tableone/index.html  

```{r}
library(tableone)
```


There are two categorical variables, sex and race. Categorical variables will need to be transformed into factors. 

```{r}
catvarstofactors <- c("sex", "race")
hw1data[catvarstofactors] <- lapply(hw1data[catvarstofactors], factor)
```


Obtain the names of the variables. 

```{r}
dput(names(hw1data))
```


Then, create a list of the variables that you want to summarize. This is a vector of variables.

```{r}
vars <-c ("age", "sex", "race", "weight", "height")
```

